---
title: "LungCapData"
author: "Amira Ibrahim"
date: "October 2, 2019"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r LungCapData, echo=TRUE}
LungCapData <- read.delim(file.choose(),header = TRUE)
attach(LungCapData)
```

#relation between Gender and Smoke :
```{r , echo=TRUE}
Table1 <- table(Gender ,Smoke)
Table1
barplot(Table1 , beside = TRUE , legend=TRUE ,xlab = "Smoking" , ylab = "Gender" ,
        main = "Relation between Gender & Smoking" ,ylim = c(0,400),col = c("blue" , "orange") )
```

#categorical variables by chisq test :
#H0 : No relation between smoking frequency and gender 
```{r , echo=TRUE}
chisq.test(Table1 , correct = TRUE)
```
##p-value > 0.05 , Fail to reject H0

#calculate OR , RR :
```{r , echo=TRUE}
library(epiR)
epi.2by2(Table1 , method = "cohort.count" , conf.level = 0.95)
```
 Odds of Females not smoking are 0.71 times odds of males not smoking
```{r , echo=TRUE}
1/0.71
```
Odds of males not smoking are 1.4 times odds of Females not smoking

##check normality 
```{r , echo=TRUE}
library(moments)
skewness(LungCap)
```
accepted level from -1 to +1 
```{r , echo=TRUE}
kurtosis(LungCap)
```
accepted level from -2 to +2 may to +3
```{r , echo=TRUE}
boxplot(LungCap)
```
visually ,data is normally distributed

###One-sample t-test for lung Capacity : 
##Test H0 =8  , conf.interval = 0.95 :
```{r , echo=TRUE}
t.test(LungCap , mu=8 , alternative = "two.sided" , conf.level = 0.95)
```
p-value >0.05 , fail to reject H0 

###Relation between Smoke & lung Capacity :
##H0 : mean of smokers = mean of non smokers :
```{r , echo=TRUE}
boxplot(LungCap~Smoke , main = "Effect of smoking on lung capacity")
```

#check variance :
```{r , echo=TRUE}
var(LungCap[Smoke == "yes"])
var(LungCap[Smoke == "no"])
```
so variance not equal  

```{r , echo=TRUE}
t.test(LungCap~Smoke , mu=0 , alternative = "two.sided" , var.eq = F, conf.level =  0.95)
```
p-value < 0.05 , reject H0 , Smoking has a significant effect on lung capacity 

#H0 : Median of lung capacity of smokers = Median of lung capacity of non smokers
```{r , echo=TRUE}
wilcox.test(LungCap~Smoke , mu=0 , alternative = "two.sided" ,
            conf.int=T,  conf.level =  0.95 , paired=F , exact=F,correct=F)
```
#p-value < 0.05 , reject H0 , Smoking has a significant effect on lung capacity
