---
title: "LungCapData"
author: "Amira Ibrahim"
date: "October 2, 2019"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r LungCapData, echo=TRUE}
LungCapData <- read.delim(file.choose(),header = TRUE)
attach(LungCapData)
```

#check names
```{r , echo=TRUE}
names(LungCapData)
```

#relation between Gender and Smoke :
```{r , echo=TRUE}
Table1 <- table(Gender ,Smoke)
Table1
barplot(Table1 , beside = TRUE , legend=TRUE ,xlab = "Smoking" , ylab = "Gender" ,
        main = "Relation between Gender & Smoking" ,ylim = c(0,400),col = c("blue" , "orange") )
```

#categorical variables by chisq test :
#H0 : No relation between smoking frequency and gender 
```{r , echo=TRUE}
chisq.test(Table1 , correct = TRUE)
```
p-value > 0.05 , Fail to reject H0

#calculate OR , RR :
```{r , echo=TRUE}
library(epiR)
epi.2by2(Table1 , method = "cohort.count" , conf.level = 0.95)
```
Odds of Females not smoking are 0.71 times odds of males not smoking
```{r , echo=TRUE}
1/0.71
```
Odds of males not smoking are 1.4 times odds of Females not smoking

##check normality 
```{r , echo=TRUE}
library(moments)
skewness(LungCap)
```
accepted level from -1 to +1 
```{r , echo=TRUE}
kurtosis(LungCap)
```
accepted level from -2 to +2 may to +3
```{r , echo=TRUE}
boxplot(LungCap)
```
visually ,data is normally distributed

###One-sample t-test for lung Capacity : 
##Test H0 =8  , conf.interval = 0.95 :
```{r , echo=TRUE}
t.test(LungCap , mu=8 , alternative = "two.sided" , conf.level = 0.95)
```
p-value >0.05 , fail to reject H0 

###Relation between Smoke & lung Capacity :
##H0 : mean of smokers = mean of non smokers :
```{r , echo=TRUE}
boxplot(LungCap~Smoke , main = "Effect of smoking on lung capacity")
```

#check variance :
```{r , echo=TRUE}
var(LungCap[Smoke == "yes"])
var(LungCap[Smoke == "no"])
```
so variance not equal  

```{r , echo=TRUE}
t.test(LungCap~Smoke , mu=0 , alternative = "two.sided" , var.eq = F, conf.level =  0.95)
```
p-value < 0.05 , reject H0 , Smoking has a significant effect on lung capacity 

#H0 : Median of lung capacity of smokers = Median of lung capacity of non smokers
```{r , echo=TRUE}
wilcox.test(LungCap~Smoke , mu=0 , alternative = "two.sided" ,
            conf.int=T,  conf.level =  0.95 , paired=F , exact=F,correct=F)
```
p-value < 0.05 , reject H0 , Smoking has a significant effect on lung capacity

#model the relation between Age , LungCap :
#use simple linear regression
```{r , echo=TRUE}
model1 <- lm(LungCap~Age)
plot(Age,LungCap,main = "Relation between Age & Lung Capacity")
abline(model1 ,col=2 , lwd=3)
```
```{r , echo=TRUE}
cor(Age,LungCap ,method="pearson")
```
there is positive strong correlation

#Denisty plots : check if the response variable is close to normal :
```{r , echo=TRUE}
library(e1071)
par(mfrow=c(1, 2))  # divide graph area in 2 columns 
plot(density(LungCap), main="Density Plot: lung capacity", ylab="Frequency")
# density plot for 'lung capacity' 
polygon(density(LungCap), col="blue") 
plot(density(Age), main="Density Plot: Age", ylab="Frequency")  # density plot for 'dist' 
polygon(density(Age), col="blue")
```

#built linear model equation :
```{r , echo=TRUE}
model1 <- lm(LungCap~Age)
model1
```
lungCap = intercept + slope*Age 
lungCap = 1.1469 + 0.5448 * Age

#check the residuals and significance
H0 : slope = 0
```{r , echo=TRUE}
summary(model1)
```
p-value <0.05 , reject H0 
67% of the variation in Lung Capicity is explained by Age

#test H0: variation mean squared regression = variation mean squared errors
```{r , echo=TRUE}
anova(model1)
sqrt(2.3)
```
p-value < 0.05 , reject H0

#Getting the coeffecient confedience interval :
```{r , echo=TRUE}
confint(model1)
```

#visualize the assumption
```{r , echo=TRUE}
plot(model1)
```

#fit a model using Age Height as explanatory variables :
H0 B0=B1=B2= 0
```{r , echo=TRUE}
mlr <- lm(LungCap~Age+Height , data = LungCapData)
summary(mlr)
```
p-value < 0.05 , reject H0
84.25% of variation in lung capacity is explained by Age and Height
increase in 1 year of Age with an increase in 0.126 of lung capacity adjusting for Height

#pearson correlation between Age ,Height :
```{r , echo=TRUE}
cor(Age , Height, method = "pearson")
```
there is +ve strong correlation

#Getting the coeffecient confedience interval :
```{r , echo=TRUE}
confint(mlr)
```

#visualize the assumption
```{r , echo=TRUE}
plot(mlr)
```

#fit model for all variables
```{r , echo=TRUE}
mlr1 <- lm(LungCap~ Age+Height+Smoke+Gender+Caesarean,data = LungCapData)
summary(mlr1)
```

#Getting the coeffecient confedience interval :
```{r , echo=TRUE}
confint(mlr1)
```

#visualize the assumption
```{r , echo=TRUE}
plot(mlr1)
```

```{r, include=FALSE}
library(tinytex)
options(tinytex.verbose = TRUE)
```

