---
title: "LungCapData"
author: "Amira Ibrahim"
date: "October 2, 2019"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r LungCapData, echo=TRUE}
LungCapData <- read.delim(file.choose(),header = TRUE)
attach(LungCapData)
```

#check names
```{r , echo=TRUE}
names(LungCapData)
```

#relation between Gender and Smoke :
```{r , echo=TRUE}
Table1 <- table(Gender ,Smoke)
Table1
barplot(Table1 , beside = TRUE , legend=TRUE ,xlab = "Smoking" , ylab = "Gender" ,
        main = "Relation between Gender & Smoking" ,ylim = c(0,400),col = c("blue" , "orange") )
```

#categorical variables by chisq test :
#H0 : No relation between smoking frequency and gender 
```{r , echo=TRUE}
chisq.test(Table1 , correct = TRUE)
```
p-value > 0.05 , Fail to reject H0

#calculate OR , RR :
```{r , echo=TRUE}
library(epiR)
epi.2by2(Table1 , method = "cohort.count" , conf.level = 0.95)
```
Odds of Females not smoking are 0.71 times odds of males not smoking
```{r , echo=TRUE}
1/0.71
```
Odds of males not smoking are 1.4 times odds of Females not smoking

##check normality 
```{r , echo=TRUE}
library(moments)
skewness(LungCap)
```
accepted level from -1 to +1 
```{r , echo=TRUE}
kurtosis(LungCap)
```
accepted level from -2 to +2 may to +3
```{r , echo=TRUE}
boxplot(LungCap)
```
visually ,data is normally distributed

###One-sample t-test for lung Capacity : 
##Test H0 =8  , conf.interval = 0.95 :
```{r , echo=TRUE}
t.test(LungCap , mu=8 , alternative = "two.sided" , conf.level = 0.95)
```
p-value >0.05 , fail to reject H0 

###Relation between Smoke & lung Capacity :
##H0 : mean of smokers = mean of non smokers :
```{r , echo=TRUE}
boxplot(LungCap~Smoke , main = "Effect of smoking on lung capacity")
```

#check variance :
```{r , echo=TRUE}
var(LungCap[Smoke == "yes"])
var(LungCap[Smoke == "no"])
```
so variance not equal  

```{r , echo=TRUE}
t.test(LungCap~Smoke , mu=0 , alternative = "two.sided" , var.eq = F, conf.level =  0.95)
```
p-value < 0.05 , reject H0 , Smoking has a significant effect on lung capacity 

#H0 : Median of lung capacity of smokers = Median of lung capacity of non smokers
```{r , echo=TRUE}
wilcox.test(LungCap~Smoke , mu=0 , alternative = "two.sided" ,
            conf.int=T,  conf.level =  0.95 , paired=F , exact=F,correct=F)
```
p-value < 0.05 , reject H0 , Smoking has a significant effect on lung capacity

#model the relation between Age , LungCap :
#use simple linear regression
```{r , echo=TRUE}
model1 <- lm(LungCap~Age)
plot(Age,LungCap,main = "Relation between Age & Lung Capacity")
abline(model1 ,col=2 , lwd=3)
```
```{r , echo=TRUE}
cor(Age,LungCap ,method="pearson")
```
there is positive strong correlation

#Denisty plots : check if the response variable is close to normal :
```{r , echo=TRUE}
library(e1071)
par(mfrow=c(1, 2))  # divide graph area in 2 columns 
plot(density(LungCap), main="Density Plot: lung capacity", ylab="Frequency")
# density plot for 'lung capacity' 
polygon(density(LungCap), col="blue") 
plot(density(Age), main="Density Plot: Age", ylab="Frequency")  # density plot for 'dist' 
polygon(density(Age), col="blue")
```

#built linear model equation :
```{r , echo=TRUE}
model1 <- lm(LungCap~Age)
model1
```
lungCap = intercept + slope*Age 
lungCap = 1.1469 + 0.5448 * Age

#check the residuals and significance
H0 : slope = 0
```{r , echo=TRUE}
summary(model1)
```
p-value <0.05 , reject H0 
67% of the variation in Lung Capicity is explained by Age

#test H0: variation mean squared regression = variation mean squared errors
```{r , echo=TRUE}
anova(model1)
sqrt(2.3)
```
p-value < 0.05 , reject H0

#Getting the coeffecient confedience interval :
```{r , echo=TRUE}
confint(model1)
```

#visualize the assumption
```{r , echo=TRUE}
plot(model1)
```

#fit a model using Age & Height as explanatory variables :
H0 B0=B1=B2= 0
```{r , echo=TRUE}
mlr <- lm(LungCap~Age+Height , data = LungCapData)
summary(mlr)
```
p-value < 0.05 , reject H0

84.25% of variation in lung capacity is explained by Age and Height

increase in 1 year of Age with an increase in 0.126 of lung capacity adjusting for Height

#pearson correlation between Age ,Height :
```{r , echo=TRUE}
cor(Age , Height, method = "pearson")
```
there is +ve strong correlation

#Getting the coeffecient confedience interval :
```{r , echo=TRUE}
confint(mlr)
```

#visualize the assumption
```{r , echo=TRUE}
plot(mlr)
```

#If we convert Height into categorical variable:
creat Height categorical A<50 , B=50-55 , c=55-60 , D=60-65 , E=65-70 , F>70 
```{r , echo=TRUE}
CatHeight <- cut(Height,breaks = c(0,50,55,60,65,70,100), labels = c("A","B","C","D","E","F"))
```
#fit model using Age , Height (as categorical variable) as explanatory variables :
```{r , echo=TRUE}
m2 <- lm(LungCap~Age+CatHeight)
summary(m2)
```
Lung capacity = 0.976+ 0.201*Age + 1.484*Xb + 2.686*Xc + 3.939*Xd + 5.007*Xe + 6.539*Xf

regression line for category A = 0.976 + 0.201*Age
regression line for category B = 2.46 + 0.201*Age
regression line for category c = 3.67 + 0.201*Age
regression line for category D = 4.92 + 0.201*Age
regression line for category E = 5.99 + 0.201*Age
regression line for category F = 7.25 + 0.201*Age

#plot Data with different colors For Height categories:
```{r , echo=TRUE}
plot(Age[CatHeight=="A"],LungCap[CatHeight=="A"] , col=2,xlim = c(0,20), ylim = c(0,15),xlab = "Age" , ylab = "Lung Capacity" , main = "Lung capacity according to Age & Height categories")
points(Age[CatHeight=="B"], LungCap[CatHeight=="B"],col=3)
points(Age[CatHeight=="C"], LungCap[CatHeight=="C"],col=4)
points(Age[CatHeight=="D"], LungCap[CatHeight=="D"],col=5)
points(Age[CatHeight=="E"], LungCap[CatHeight=="E"],col=6)
points(Age[CatHeight=="F"], LungCap[CatHeight=="F"],col=7)
legend(0,15.5,legend = c("A","B","C","D","E","F"),col = 2:7,pch = 1,cex = 0.8)
abline(a=0.976,b=0.201,col=2,lwd=3)
abline(a=2.46,b=0.201,col=3,lwd=3)
abline(a=3.67,b=0.201,col=4,lwd=3)
abline(a=4.92,b=0.201,col=5,lwd=3)
abline(a=5.99,b=0.201,col=6,lwd=3)
abline(a=7.25,b=0.201,col=7,lwd=3)
```
increase in 1 year associate with 0.201 change in lung capacity independent on Height categories

Age effect is the same for all Height categories

#visualize the assumption
```{r , echo=TRUE}
plot(m2)
```

#fit model using Age , Smoking as explanatory variables :
```{r , echo=TRUE}
mlr1 <- lm(LungCap~Age+Smoke)
summary(mlr1)
```
67% of variation in lung capacity is explained by Age & Smoke 

increase in 1 year associate with 0.555 change in lung capacity in non smokers

LungCap = 1.086 + (0.555*Age) +(-0.649*Smoke yes)

#Plot the data to differ between Smokers & non smokers :
```{r , echo=TRUE}
plot(Age[Smoke=="no"],LungCap[Smoke=="no"] , col="blue" , ylim = c(1,15),xlab = "Age" , ylab = "Lung Capacity" , main = "Lung capacity according to Age & Smoke")
points(Age[Smoke=="yes"], LungCap[Smoke=="yes"],col="red",pch=16)
legend(3,15,legend = c("Non Smokers","Smokers"),col = c("blue","red"),pch = c(1,16),bty = "n")
abline(a=1.08,b=0.555,col="blue",lwd=3)
abline(a=0.431,b=0.555,col="red",lwd=3)
```
increase in 1 year associate with 0.555 change in mean lung capacity , this increase is the same in Smokers & non Smokers 

For Smokers mean lung capacity decreased by 0.649 ,this decrease is the same in All ages 

Effect of Age is independent on Smoking & vise versa , So no interaction between Age and Smoke

#Getting the coeffecient confedience interval :
```{r , echo=TRUE}
confint(mlr1)
```

#visualize the assumption
```{r , echo=TRUE}
plot(mlr1)
```

#fit model for all variables
```{r , echo=TRUE}
mlr2 <- lm(LungCap~ Age+Height+Smoke+Gender+Caesarean,data = LungCapData)
summary(mlr2)
```
Lung capacity = -11.32+(0.16*Age) + (0.26*Height)+(-0.061*Smoke yes)+(0.38*Gender male)+(-0.21*caesarean yes)

85.32% of variation in Lung capacity is explained by other variables

#Getting the coeffecient confedience interval :
```{r , echo=TRUE}
confint(mlr2)
```

#visualize the assumption
```{r , echo=TRUE}
plot(mlr2)
```

```{r, include=FALSE}
library(tinytex)
options(tinytex.verbose = TRUE)
```

